Designing a high-performance boundary element library with OpenCL and Numba

### 2021-04-14 bem.tex
Finished small corrections.

Once the grid, the space objects, and the operator(s) are defined, the main computational step is performed, namely the \textbf{computation of the discrete matrix entries} $\dmat{A}_{ij}$. For pairs of triangles $\tau_i$ and $\tau_j$ that do not share a joint edge or vertex this is done through a simple numerical quadrature rule that approximates $\dmat{A}_{ij}\approx \sum_{\ell}\sum_{q} g(\bx_\ell, \by_q)\psi_i(\bx_\ell)\phi_j(\by_q)\omega_\ell\omega_q$, where the $\bx_\ell$ and $\by_q$ are quadrature points in the corresponding triangles $\tau_i$ and $\tau_j$, and the values $\omega_i$ and $\omega_j$ are the quadrature weights. In the case that two triangles share a joint vertex/edge or the triangles $\tau_i$ and $\tau_j$ are identical, corresponding singularity adapted quadrature numerical quadrature rules are used that are based on singularity removing coordinate transformations.

The values $\dvec{b}_i$ of the right-hand side vector $b$ are similarly computed through a numerical quadrature rule.

In the final step, \textbf{Bempp solves the underlying linear system of equations} either through a direct LU decomposition or through iterative solvers such as GMRES. The solution can then be evaluated away from the surface $\Gamma$ through domain potential operators and exported in various formats for visualization.

In summary, to solve a boundary integral equation problem, the following steps are performed by Bempp:
\begin{enumerate}
	\item Import of the surface description as triangulation data.
	\item Definition of the trial space, test space and range space.
	\item Discretization into a matrix problem $\dmat{A}\dvec{x}=\dvec{b}$.
	\item Solution of the matrix problem by either a direct or iterative solver.
	\item Evaluation of domain potential operators for visualization and post-processing.
\end{enumerate}

>>>>

Once the grid, the space objects, and the operator(s) are defined, the main computational step is performed, namely the \textbf{computation of the discrete matrix entries} $\dmat{A}_{ij}$. For pairs of triangles $\tau_i$ and $\tau_j$ that do not share a joint edge or vertex this is done through a simple numerical quadrature rule that approximates $\dmat{A}_{ij}\approx \sum_{\ell}\sum_{q} g(\bx_\ell, \by_q)\psi_i(\bx_\ell)\phi_j(\by_q)\omega_\ell\omega_q$, where the $\bx_\ell$ and $\by_q$ are quadrature points in the corresponding triangles $\tau_i$ and $\tau_j$, and the values $\omega_\ell$ and $\omega_q$ are the quadrature weights. In the case that two triangles share a joint vertex/edge or the triangles $\tau_i$ and $\tau_j$ are identical (see Figure \ref{fig:triangles}), corresponding singularity adapted quadrature numerical quadrature rules are used that are based on singularity removing coordinate transformations.

The values $\dvec{b}_i$ of the right-hand side vector $\dvec{b}$ are similarly computed through a numerical quadrature rule.

In the final step, \textbf{Bempp-cl solves the underlying linear system of equations} either through a direct LU decomposition or through iterative solvers. The solution can then be evaluated away from the surface $\Gamma$ through domain potential operators and exported in various formats for visualization.

In summary, to solve a boundary integral equation problem, the following steps are performed by Bempp-cl:
\begin{enumerate}
	\item Import of the surface description as triangulation data.
	\item Definition of the trial space, test space and range space.
	\item Discretization into a matrix problem $\dmat{A}\dvec{x}=\dvec{b}$.
	\item Solution of the matrix problem by either a direct or iterative solver.
	\item Evaluation of domain potential operators for visualization and post-processing.
\end{enumerate}

### 2021-04-14 bempp_overview.tex
Finished small corrections.

The main computational cost involved in solving a problem using boundary element methods is due to discretising the boundary integral operator on the left-hand side of \cref{eq:bnd_integral} to obtain the matrix $\dmat{A}$: using dense methods, discretising an operator has quadratic complexity in terms of the number of surface triangles. For larger problems this cost can be reduced through the use of approximate hierarchical matrices or fast multipole methods (FMM), where the price of the improved complexity is significantly more involved data structures and additional approximation errors. Bempp-cl provides interfaces to Exafmm \cite{bempp_exafmm} for large problems. Here, we focus on dense discretisation that is natively implemented in Bempp-cl and is suitable for medium sized problems up to a few ten thousand elements, depending on available memory. Great care needs to be taken to ensure that the quadratic complexity operator assembly routines perform as highly as possible.

@@

For OpenCL assembly, the code checks additional parameters, such as the default vector length for SIMD operators (e.g., 4 for double precision and 8 for single precision in Intel AVX2, or 1 if a GPU is used), and whether the discretization should proceed in single or double precision. The OpenCL kernel is then compiled for the underlying compute device using PyOpenCL and executed. If the computational backend is Numba, the call is forwarded to the corresponding Numba discretization routines and executed.

>>>>

The main computational cost involved in solving a problem using boundary element methods is due to discretising the boundary integral operator on the left-hand side of \cref{eq:bnd_integral} to obtain the matrix $\dmat{A}$: using dense methods, discretising an operator has quadratic complexity in terms of the number of surface triangles. For larger problems this cost can be reduced through approximation techniques, such as hierarchical (H-) matrices or fast multipole methods (FMM) with log-linear or even linear complexity. The price of the improved complexity is significantly more involved data structures and additional approximation errors. Bempp-cl provides interfaces to Exafmm \cite{bempp_exafmm} for large problems. Here, we focus on dense discretisation that is natively implemented in Bempp-cl and is suitable for medium sized problems up to a few ten thousand elements, depending on available memory. Great care needs to be taken to ensure that the quadratic complexity operator assembly routines perform as highly as possible.

@@

For OpenCL assembly, the code checks additional parameters, such as the default vector length for SIMD operators (e.g., 4 for double precision and 8 for single precision in Intel AVX, or 1 if a GPU is used), and whether the discretization should proceed in single or double precision. The OpenCL kernel is then compiled for the underlying compute device using PyOpenCL and executed. If the computational backend is Numba, the call is forwarded to the corresponding Numba discretization routines and executed.

### 2021-04-14 intro.tex
Finished small corrections.

\chapterinitial{The Bempp boundary element library} (originally named BEM++) started in 2011 as a project to develop an open-source C++ library for the fast solution of boundary integral equations. The original release came with a simple Python wrapper to the C++ library \cite{bempp_orig}. Over time, more and more functionality was moved into the Python interface, while computationally intensive routines and the main data structures remained in C++.

At the end of 2019, we completed the main steps of a full rewrite of Bempp and released the first version (0.1) of Bempp-cl \cite{Bempp-cl}. This was followed later in 2020 by version 0.2, the first release that we considered feature complete and mature for application use. Since then we have used Bempp-cl in a number of practical applications and many of our users are migrating to it from the old C++ based Bempp. In this article, we discuss the motivation for the rewrite and restructure of Bempp and the reasoning behind the design choices we made when writing Bempp-cl.

@@

We begin this article by giving an overview of the boundary element method, and looking at how Bempp can be used to implement such problems. Following this, we discuss the implementation of boundary element kernels using OpenCL in more detail, and provide a number of performance benchmarks on different compute devices, including CPUs, AMD, Intel, and Nvidia GPUs. We conclude with some thoughts on the advantages and disadvantages of OpenCL and Numba.

>>>>

\chapterinitial{The Bempp boundary element library} (originally named BEM++) started in 2011 as a project to develop an open-source C++ library for the fast solution of boundary integral equations. The original release came with a simple Python wrapper to the C++ library \cite{bempp_orig}. Over time, more and more functionality was moved into the Python interface, while computationally intensive routines and the main data structures remained in C++.
At the end of 2019, we completed the main steps of a full rewrite of Bempp and released the first version (0.1) of Bempp-cl . This was followed later in 2020 by version 0.2, the first release that we considered feature complete and mature for application use \cite{Bempp-cl}. Since then we have used Bempp-cl in a number of practical applications and many of our users are migrating to it from the old C++ based Bempp. In this article, we discuss the motivation for the rewrite and restructure of Bempp and the reasoning behind the design choices we made when writing Bempp-cl.

@@

We begin this article by giving an overview of the boundary element method, and looking at how Bempp can be used to implement such problems. Following this, we discuss the implementation of boundary element kernels using OpenCL in more detail, and provide a number of performance benchmarks on different compute devices, including CPUs and Nvidia GPUs. We conclude with some thoughts on the advantages and disadvantages of OpenCL and Numba.

### 2021-04-14 opencl.tex
Finished small corrections.

OpenCL \cite{opencl} is a heterogeneous compute standard for CPUs, GPUs, FPGAs, and other types of devices that provide conformant drivers. At its core, OpenCL executes compute kernels that can be written in C99, or (more recently) in C++. The current version of OpenCL is 3.0, though the most widely implemented standard is OpenCL 1.2, which Bempp-cl uses.

OpenCL splits computational tasks into work-items, each of which represents a single execution of a compute kernel. Work items are grouped together into work groups, which share local memory. Barrier synchronization is only allowed within a work group. All work items are uniquely indexed by a one, two, or three dimensional index space, called NDRange. Kernels are launched onto a compute device (e.g., a CPU or GPU) from a host device. OpenCL allows kernels to be loaded as strings and compiled on the fly for a given device, making it well suited for launching from high-productivity languages.

To launch an OpenCL kernel the user must first initialize buffers on the compute device, then copy any relevant data from host to the compute device. A kernel string can then be loaded and just-in-time compiled for the device. The kernel is then run, and the results can be copied back to the host.

@@

Each work-item computes (using numerical quadrature) all interactions of basis functions on the trial element with basis functions on the test element. Before summing the result into the global result buffer, the kernel checks via the connectivity information if the test and trial triangles are adjacent or identical (see \cref{fig:triangles}). To do this, it simply checks if at at least one of the node indices of the test triangle is equal to one of the node indices of the trial triangle. If this is true and the grids are not disjoint, the result of the kernel is discarded and not summed back into the global result buffer: for these triangles, separate singular quadrature rules need to be used. The effect is that a few work-items do work that is discarded. However, in a grid with $N$ elements, the number of triangle pairs requiring a singular quadrature rule is $\mathcal{O}(N)$, while the total number of interactions is $N^2$. Hence, only a tiny fraction of work-items are discarded.

@@

For this assembly, we create one workgroup for each triangle pair. Inside this workgroup, we have a number of work-items that evaluate the quadrature rules then sum up the results. Depending on how two triangles are related to each other, different types of singular quadrature rule are needed. We solve this by pre-loading all possible quadrature rules onto the device, and also store for each triangle pair an index pointing to the required quadrature rule so that the kernel function can select the correct rule to evaluate. For the singular quadrature rules, we did not implement separate SIMD optimized kernels as the proposed implementation is already highly efficient and requires only a fraction of the computational time of the regular quadrature rules described above. At the end, the singular integral values are either summed into the overall result matrix, or (if desired by the user) stored as separate sparse matrix.

>>>>

OpenCL \cite{opencl} is a heterogeneous compute standard for CPUs, GPUs, FPGAs, and other types of devices that provide conformant drivers. At its core, OpenCL executes compute kernels that can be written in C99, or (more recently) in C++, with some restrictions on the allowed operations. The current version of OpenCL is 3.0, though the most widely implemented standard is OpenCL 1.2, which Bempp-cl uses.

OpenCL splits computational tasks into work-items, each of which represents a single execution of a compute kernel. Work-items are grouped together into work-groups, which share local memory. Barrier synchronization is only allowed within a work-group. All work items are uniquely indexed by a one, two, or three dimensional index space, called NDRange. Kernels are launched onto a compute device (e.g., a CPU or GPU) from a host device. OpenCL allows kernels to be loaded as strings and compiled on-the-fly for a given device, making it well suited for launching from high-productivity languages.

To launch an OpenCL kernel the user must provide relevant data as buffers, which are transferred from host to the corresponding compute device. A kernel string can then be loaded and just-in-time compiled for the device. The kernel is then run, and the results can be copied back to the host.

@@

Each work-item computes (using numerical quadrature) all interactions of basis functions on the trial element with basis functions on the test element. Before summing the result into the global result buffer, the kernel checks via the connectivity information if the test and trial triangles are adjacent or identical (see \cref{fig:triangles}). To do this, it simply checks if at at least one of the node indices of the test triangle is equal to one of the node indices of the trial triangle. If this is true and the grids are not disjoint, the result of the kernel is discarded and not summed back into the global result buffer: for these triangles, separate singular quadrature rules need to be used. The effect is that a few work-items do work that is discarded. However, in a grid with $N$ elements, the number of triangle pairs requiring a singular quadrature rule is $\mathcal{O}(N)$, while the total number of triangle interactions is $N^2$. Hence, only a tiny fraction of work-items are discarded.

@@

For this assembly, we create one work-group for each singular triangle pair. Inside this work-group, we have a number of work-items that evaluate the quadrature rules then sum up the results. Depending on how two triangles are related to each other, different types of singular quadrature rule are needed. We solve this by pre-loading all possible quadrature rules onto the device, and also store for each triangle pair an index pointing to the required quadrature rule so that the kernel function can select the correct rule to evaluate. For the singular quadrature rules, we did not implement separate SIMD optimized kernels as the proposed implementation is already highly efficient and requires only a fraction of the computational time of the regular quadrature rules described above. At the end, the singular integral values are either summed into the overall result matrix, or (if desired by the user) stored as separate sparse matrix.

### 2021-04-14 performance.tex
Finished small corrections.

In this section, we provide a number of performance benchmarks. The tests are running on a Dell Precision 7740 Workstation Laptop with 64 GB RAM. Its CPU is an Intel i9-9980HK with a base clock of 2.4GHz and a burst clock of 5GHz. The CPU supports AVX, AVX2, and AVX-512.
As a GPU we use an Nvidia Quadro RTX 3000 GPU. All benchmark tests were performed in Linux. For OpenCL on the GPU, we use the Nvidia GPU drivers and as CPU runtime we compare the open-source PoCL driver against the Intel CPU OpenCL runtime environment. All timing runs were repeated several times to make sure that the overhead from running the OpenCL and Numba JIT compilers did not skew the results. Though hardly noticeable by users, for smaller experiments the JIT phase typically takes longer than the actual computation.

@@

We first compare the OpenCL CPU performance for the Intel OpenCL runtime, and the PoCL OpenCL runtime driver. We run the tests in single precision and double precision. The native vector width for both drivers in single precision is $8$, and in double precision is $4$, which corresponds to AVX instructions. In Bempp-cl, this means that we assemble in vectorized form one test triangle with $8$ trial triangles in single precision, and with $4$ trial triangles in double precision. Within the assembly, almost all floating point instructions are manually vectorized to take advantage of this. We should hence see up a factor 2 speed-up between single and double precision assembly.

@@

\begin{figure}
\begin{center}
\begin{tabular}{l|c|c}
	        &   single      &    double\\
	        \hline
	 PoCL   &   0.05s       &    0.08s\\
	 Numba  &   0.25s       &    0.25s\\
	 GPU    &   0.11s       &    3.00s\\
\end{tabular}
\end{center}
\caption{Comparison of PoCL, Numba and GPU Assembly for a grid with 2048 elements.}
\label{fig:cpu_gpu_numba_compare}
\end{figure}
In Bempp-cl, we can easily switch between CPU Assembly, GPU Assembly, and Numba Assembly. \Cref{fig:cpu_gpu_numba_compare} shows a comparison between these modes for a grid with 2048 elements. The speed differences are striking. GPU assembly in single precision is a factor of two slower than CPU assembly (and much slower in double precision due to Nvidia hardware restrictions for double precision). Numba is five times slower for single precision and still around three times slower for double precision than PoCL. The GPU behaviour can be explained by data transfer: while the GPU kernels themselves are extremely fast, data transfer over the bus severely limits performance. Even for medium sized problems, we have to transfer the data back to the main memory as GPU RAM is too limited to keep dense matrices with a tens of thousands or rows and columns on the device.

An alternative method is to compute the matrix-vector product (matvec) $\dmat{A}\dvec{x}$ on the device without first computing the dense matrix $\dmat{A}$. This can be done by recomputing all matrix elements during each matvec calculation on-the-fly and not storing them. We have done experiments with this, and observed significant speed-ups compared to CPU evaluation as we now only need to transfer single vectors over the bus. For larger problems, however, it is not competitive compared to accelerated methods such as fast multipole methods (FMM), due to the quadratic complexity of direct evaluation of the matvec compared to linear complexity of FMM. For smaller problems, it is still practically better to just assemble the whole matrix and store it, as then matvecs are much faster for an iterative solver. Hence, we can conclude that there is only limited practical relevance for on-the-fly GPU evaluation of boundary integral operators. There are, however, very significant practical advantages of on-the-fly evaluation of the evaluation of domain potential operators for visualization and post-processing, as we will see in the next section.

The Numba performance difference seen in \cref{fig:cpu_gpu_numba_compare} is interesting. The main reason for this, we believe, is significantly lower usage of AVX vectorized instructions. We have taken care to optimize the Green's function evaluation for auto-vectorized evaluation in Numba, but the loops over integration points and other operations auto-vectorize very badly when just looping over all trial triangle for a single test triangle, as we currently do. We could tune our code for better auto-vectorization in Numba, but this would give little benefit as we have highly optimized hand-tuned OpenCL kernels already. We therefore recommend using Numba for operator assembly only as a fallback if no OpenCL runtime is available.

@@

In this section, we consider the evaluation of the electric potential operator, defined by
\begin{multline}
\left(\mathcal{E}\mathbf{p}\right)(\bx) = \ii k\int_{\Gamma}\mathbf{p}(\by)g(\bx, \by) \ds[\by]\\
-\frac{1}{\ii k}\nabla_{\bx}\int_{\Gamma}\operatorname{div}\mathbf{p}(\by)g(\bx, \by)\ds[\by],
\end{multline}
where $g(\bx, \by) = \frac{\ee^{\ii k|\bx - \by|}}{4\pi|\bx - \by|}$ is the Helmholtz Green's function, and
$k$ the wavenumber of the problem.
The function $\mathbf{p}:\mathbb{R}^3\to\mathbb{R}^3$ is a vector-valued function, leading to an overall vector solution at each point $\bx$. The implementation of electromagnetic problems in Bempp is covered in detail in \cite{bempp_maxwell}. For these experiments, we again use a grid with 32,768 elements and RWG (Rao, Wilton, and Glisson) edge-based basis functions. For the potential evaluation, we use 50,000 random evaluation points in the exterior of the unit sphere, and as wavenumber use $k=1.0$.

In \cref{fig:efield_domain_potential}, we compare the performance of GPU evaluation with that of the PoCL CPU driver. In single-precision, the GPU significantly outperforms the CPU; and even in double-precision, the Nvidia Quadro RTX GPU is faster than the 8-core CPU, even though its hardware is not optimised for fast double-precision operations.

@@

\section{SUMMARY}
With Bempp-cl, we have created a Python library that achieves high-performance through use of modern just-in-time compilation technologies. Bempp-cl mixes Numba evaluation for less compute intensive linear complexity loops and sparse matrix generation with highly optimized OpenCL kernels for computationally sensitive dense matrix assembly routines. Basing development in Python instead of classical C/C++ makes it very easy to adapt the library and integrate with other libraries with Python interfaces, such as the recent work integrating Bempp-cl and the Exafmm library to build electrostatic virus-scale simulations \cite{bempp_exafmm}.

Strict separation of the computational backend from the interface layer in the library also makes it easy to integrate further backends, allowing us to remain up to date with future compute models. OpenCL itself has proved a valuable choice for the purpose of this library, as it allows us to run CPU and GPU optimized kernels with very little code overhead, and allows the user to easily move between CPU and GPU based compute devices with a simple parameter change. In this article, we demonstrated Nvidia benchmarks, the same benchmarking code could be used to run on AMD or Intel GPUs (for example).

@@

Overall, with the model of mixed Python/OpenCL/Numba development, we have created a flexible and easy to extend platform for integral equation computations. The initial re-implementation efforts are really paying off, as they allow us to develop new features in a far simpler environment without sacrificing performance. Strict separation of compute backends and higher level routines makes it easy for us to integrate other accelerator techniques in the future with little code changes, and to react to new trends in heterogeneous computing.

>>>>

In this section, we provide a number of performance benchmarks. The tests are running on a Dell Precision 7740 Workstation Laptop with 64 GB RAM. Its CPU is an Intel i9-9980HK with a base clock of 2.4GHz and a burst clock of 5GHz. The CPU supports AVX, AVX2, and AVX-512.
As a GPU we use an Nvidia Quadro RTX 3000 GPU. All benchmark tests were performed in Linux. For OpenCL on the GPU, we use the Nvidia GPU drivers and as CPU runtime we compare the open-source PoCL driver against the Intel CPU OpenCL runtime environment. All timing runs were repeated several times to make sure that the overhead from running the OpenCL and Numba just-in-time compilers did not skew the results. Though hardly noticeable by users, for smaller experiments the compilation phase typically takes longer than the actual computation.

@@

We first compare the OpenCL CPU performance for the Intel OpenCL runtime, and the PoCL OpenCL runtime driver. We run the tests in single precision and double precision. The native vector width for both drivers in single precision is $8$, and in double precision is $4$, corresponding to AVX instructions. In Bempp-cl, this means that we assemble in vectorized form one test triangle with $8$ trial triangles in single precision, and with $4$ trial triangles in double precision. Within the assembly, almost all floating point instructions are manually vectorized to take advantage of this. We should hence see up to a factor 2 speed-up between single and double precision assembly.

@@

\begin{figure}
\begin{center}
\begin{tabular}{l|c|c}
	        &   single      &    double\\
	        \hline
	 PoCL   &   0.05s       &    0.08s\\
	 Numba  &   0.25s       &    0.25s\\
	 GPU    &   0.11s       &    3.00s\\
\end{tabular}
\end{center}
\caption{Comparison of PoCL, Numba and GPU Assembly for a grid with 2048 elements.}
\label{fig:cpu_gpu_numba_compare}
\end{figure}
In Bempp-cl, we can easily switch between CPU Assembly, GPU Assembly, and Numba Assembly. \Cref{fig:cpu_gpu_numba_compare} shows a comparison between these modes for a grid with 2048 elements. The speed differences are striking. GPU assembly in single precision is a factor of two slower than CPU assembly (and much slower in double precision due to Nvidia hardware restrictions for double precision). Numba is five times slower for single precision and still around three times slower for double precision than PoCL. The GPU behaviour can be explained by data transfer: while the GPU kernels themselves are extremely fast, data transfer over the bus severely limits performance. Even for medium sized problems, we have to transfer the data back to the main memory as GPU RAM is too limited to keep dense matrices with tens of thousands of rows and columns on the device.

An alternative method is to compute the matrix-vector product (matvec) $\dmat{A}\dvec{x}$ on the device without first computing the dense matrix $\dmat{A}$. This can be done by recomputing all matrix elements during each matvec calculation on-the-fly and not storing them. We have done experiments with this, and observed significant speed-ups compared to CPU evaluation as we now only need to transfer single vectors over the bus. For larger problems, however, it is not competitive compared to accelerated methods such as FMM, due to the quadratic complexity of direct evaluation of the matvec compared to linear complexity of FMM. For smaller problems, it is still practically better to just assemble the whole matrix and store it, as then matvecs are much faster for an iterative solver. Hence, we can conclude that there is only limited practical relevance for on-the-fly GPU evaluation of boundary integral operators. There are, however, very significant practical advantages of on-the-fly evaluation of domain potential operators for visualization and post-processing, as we will see in the next section.

The Numba performance difference seen in \cref{fig:cpu_gpu_numba_compare} is interesting. The main reason for this, we believe, is significantly lower usage of AVX vectorized instructions. We have taken care to optimize the Green's function evaluation for auto-vectorized evaluation in Numba, but the loops over integration points and other operations auto-vectorize very badly when just looping over all trial triangles for a single test triangle, as we currently do. We could tune our code for better auto-vectorization in Numba, but this would give little benefit as we have highly optimized hand-tuned OpenCL kernels already. We therefore recommend using Numba for operator assembly only as a fallback if no OpenCL runtime is available.

@@

In this section, we consider the evaluation of the electric potential operator, defined by
\begin{multline}
\left(\mathcal{E}\mathbf{p}\right)(\bx) = \ii k\int_{\Gamma}\mathbf{p}(\by)g(\bx, \by) \ds[\by]\\
-\frac{1}{\ii k}\nabla_{\bx}\int_{\Gamma}\operatorname{div}\mathbf{p}(\by)g(\bx, \by)\ds[\by],
\end{multline}
where $g(\bx, \by) = \frac{\ee^{\ii k|\bx - \by|}}{4\pi|\bx - \by|}$ is the Helmholtz Green's function, and
$k$ the wavenumber of the problem.
The function $\mathbf{p}:\mathbb{R}^3\to\mathbb{R}^3$ is a vector-valued function, leading to an overall vector solution at each point $\bx$. The implementation of electromagnetic problems in Bempp is covered in detail in \cite{bempp_maxwell}. For these experiments, we again use a grid with 32,768 triangles but this time RWG (Rao, Wilton, and Glisson) edge-based basis functions. For the potential evaluation, we use 50,000 random evaluation points in the exterior of the unit sphere, and as wavenumber use $k=1.0$.

In \cref{fig:efield_domain_potential}, we compare the performance of GPU evaluation with that of the PoCL CPU driver. In single-precision, the GPU significantly outperforms the CPU; and even in double-precision, the Nvidia Quadro RTX GPU is faster than the 8-core CPU, even though its hardware is not optimised for fast double precision operations.

@@

\section{SUMMARY}
With Bempp-cl, we have created a Python library that achieves high-performance through use of modern just-in-time compilation technologies. Bempp-cl mixes Numba evaluation for less compute intensive linear complexity loops and sparse matrix generation with highly optimized OpenCL kernels for computationally sensitive dense matrix assembly routines. Basing development on Python instead of classical C/C++ makes it very easy to adapt the library and integrate with other libraries with Python interfaces, such as the recent work integrating Bempp-cl and the Exafmm library to build electrostatic virus-scale simulations \cite{bempp_exafmm}.

Strict separation of the computational backend from the interface layer in the library also makes it easy to integrate further backends, allowing us to remain up to date with future compute models. OpenCL itself has proved a valuable choice for the purpose of this library, as it allows us to run CPU and GPU optimized kernels with very little code overhead, and allows the user to easily move between CPU and GPU based compute devices with a simple parameter change. In this article, we demonstrate Nvidia benchmarks, the same benchmarking code could be used to run on AMD or Intel GPUs.

@@

Overall, with the model of mixed Python/OpenCL/Numba development, we have created a flexible and easy to extend platform for integral equation computations. The initial re-implementation efforts by abandoning our old C++ code base are paying off, as they allow us to develop new features in a far simpler environment without sacrificing performance. Strict separation of compute backends and higher level routines makes it easy for us to integrate other accelerator techniques in the future with little code changes, and to react to new trends in heterogeneous computing.
